---
title: "California Spiny Lobster Abundance"
author: "Zoe Duerksen-Salm and Audrey Nolan"
date: "11/18/2019"
output: html_document
---
Thoughts: 
- move all of the figure caption/blurbs to before the graphs. Allison did this in her example report and i like the look of it. 
- How do we find a proper citation? I only found DOI. Do we need to include date data retrieved etc?
- for discussion of results from section C, do we want to remove the questions and just make a separate paragraph for each test, then talk about the main graph? Orrrrr talk about the main graph, and throw in the specific statitic values through the discussion. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

library(tidyverse)
library(tidyr)
library(janitor)
library(here)
library(plotly)
library(DT)
library(lubridate)
library(ggridges)
library(knitr)
library(kableExtra)

```

### Introduction 
A short (7 - 8 sentence) introduction that helps the reader understand the purpose/motivation of the research, and previews the content of the report.

This report explores California Spiny Lobster, *Panulirus interruptus*, populations from 2012-2018 at five sites off the coast of Santa Barbara, California. Spiny lobsters are play an important role in the giant kelp forests by foraging on algae, snails and crabs. Additionally, spiny lobsters feed on sea urchins which helps maintain the health of the kelp forest. Two of the study sites, Naples and Isla Vista, became marine protected areas(MPA) in 2012. Under the Marine Life Protection Act, MPAs aim to protect the diversity and abundance of marine life, the habitats they depend on, and the integrity of marine ecosystems(Cal Department of Fish and Wildlife). MPAs allow for.... 

### Data and Methods
A brief ‘Data and Methods’ section (5 - 6 sentences) summarizing the data and how it was analyzed. Briefly describe the variables being studied in this report. Include the types of statistical tests performed, significance level used, and software (with version) for analysis.

---

The data used for this report was obtained by reasearchers with Santa Barbara Coastal Long Term Ecological Research(SBC LTER) that began sampling in 2012 and [remains] ongoing. The number of lobsters and their sizes are recorded by divers in four 300m^2 lobster transects. This data is collected annually in late summer at each of the five sites along the coast of the Santa Barbara Channel. The size of lobsters(mm) and count were the two main varibles used within this report. 

Statistical tests were performed on the data including a Whelch Two Sample t-test with a significance level of p = 0.05. All statistical analyses and figures were prepared using R software version 3.6.1.

doi:10.6073/pasta/a593a675d644fdefb736750b291579a0

```{r}
#Read in lobster file + make it tidy

lobster_raw <- read_csv("lobster_abundance_sbc_lter.csv", na = "-99999")

lobster_tidy <- lobster_raw %>% 
  clean_names() %>%
  tidyr::uncount(count) %>% 
  select(year:size_mm)



```

### Results

Prepare finalized data visualizations and perform statistical analyses (as indicated) to do the following. You are expected to also add useful text that weaves the pieces together into a cohesive Results section (i.e., this should not just be a series of figures and without useful description & text flow between them). 

#### A: Changes in Lobster abundance (counts) by site.

In text (above or below the graph for Results A), describe general / interesting trends that you observe for lobster abundance at the 5 sites.

---

MPA sites established in 2012 - Isla Vista and Naples - are represented by orange coloration in all figures throughout this report. Non-MPA sites - Arroyo Quemado, Carpinteria, Mohawk - are indicated with blue coloration.

```{r}
# creating dataframe labeling MPA vs nonMPA
lobster_abund <- lobster_tidy %>% 
  mutate(protection = ifelse(site %in% c("IVEE", "NAPL"), "MPA", "non-MPA"))

# Change site variable names from abreviations to full spelling
lobster_abund$site <- gsub("IVEE", "Isla Vista", lobster_abund$site)
lobster_abund$site <- gsub("NAPL", "Naples", lobster_abund$site)
lobster_abund$site <- gsub("AQUE", "Arroyo Quemado", lobster_abund$site)
lobster_abund$site <- gsub("CARP", "Carpinteria", lobster_abund$site)
lobster_abund$site <- gsub("MOHK", "Mohawk", lobster_abund$site)

```

```{r}
# Making site abundance graph

site_abundance_graph <- ggplot(lobster_abund, aes(x = year))+
  geom_histogram(binwidth = 1,
                 aes(fill = protection),
                 show.legend = FALSE) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9"))+
  facet_wrap(~site) +
  scale_x_continuous(limits = c(2011, 2019),
                     expand = c(0,0),
                     breaks = c(2012, 2014, 2016, 2018))+
  theme_light() +
  labs(x = "Year",
       y = "Lobster Count (n)",
       title = "Fig. 1: Lobster Abundance Across Five Sites", 
       caption = "Figure 1 depicts the lobster abundance at each of the five sites. MPA sites are\n represented by orange coloration and non-MPA sites are shown by blue coloration.")

site_abundance_graph

```
Figure one [shows] the differences in lobster abundance over a six-year period. Naples and Isla Vista, the two MPAs, [show] a steady increase in lobster abundance during the period of study. The non-MPA sites, Arroyo Quemada, Carpinteria, and Mohawk, show varying lobster abundance over the six-year period. Mohawk shows a decrease in abundance after the highest count at the site in 2014. This steady decline in abundance may be due to overharvesting during the peak abundance year and the lobster populations have [struggled] to recover. 


#### B: Differences in lobster size distribution from 2012 to 2018

Visually explore lobster size distribution shifts
Compare lobster sizes in 2012 and 2018

```{r}

lobster_12_18 <- lobster_tidy %>% 
  filter(year %in% c(2012, 2018))

lobster_all <- lobster_abund %>% 
  filter(year %in% c(2012, 2018))

lobster_size <- lobster_12_18 %>% 
  select(year, site, size_mm)

# Change site variable names from abreviations to full spelling
lobster_size$site <- gsub("IVEE", "Isla Vista", lobster_size$site)
lobster_size$site <- gsub("NAPL", "Naples", lobster_size$site)
lobster_size$site <- gsub("AQUE", "Arroyo Quemado", lobster_size$site)
lobster_size$site <- gsub("CARP", "Carpinteria", lobster_size$site)
lobster_size$site <- gsub("MOHK", "Mohawk", lobster_size$site)

```

```{r}
# Creating a graph to compare lobsters sizes in 2012 vs. 2018 

size_dist_by_site <- ggplot(lobster_all, aes(x = size_mm, y = site)) +
  geom_density_ridges(aes(fill = protection),
                      show.legend = FALSE)+
  scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
  facet_wrap(~year) +
  scale_x_continuous(limits = c(30, 135),
                     expand = c(0,0)) +
  theme_minimal()+
  labs(x = "Lobster Size (mm)",
       y = "Research Site",
       title = "Fig. 2: Lobster Size Across Sites",
       caption = "Figure 2 illustrates the difference in lobsters size between 2012 and 2018 at all five sites.")

size_dist_by_site
```

In text, describe general / interesting trends that you see in community size structure at different sites (again, considering MPA vs. non-MPA status) for lobsters between 2012 and 2018.

In 2012 when Naples and Isla Vista were designated as MPAs, the sites had relatively smaller sized lobsters across the distrubution of sizes. However, the 2018 data indicates that the populations at these sites shifted towards larger lobsters, likley due to the designation of this area as an MPA, allowing for lobsters to mature without being harvested. Non-MPA sites show minimal change in the distrubution of lobster sizes over the six-year period. 

#### C: Comparing mean lobster size at MPA vs. non-MPA sites in 2012 and 2018 

```{r}
# For each year (2012 and 2018) consolidate the size observations into only two groups: MPA and non-MPA lobsters.

# Manipulate df to indicate MPA or non-MPA status. Filter df to include only 2012 + 2018 observations.

lobster_protection <- lobster_tidy %>% 
  mutate(protection = ifelse(site %in% c("IVEE", "NAPL"), "MPA", "non-MPA")) %>% 
  filter(year == c(2012, 2018)) %>% 
  select("year", "site", "size_mm", "protection")

#### wrangling to summarize mean of each 4 groups 

# MPA 2018
lobster_mean_MPA_2018 <- lobster_protection %>% 
  filter(year == 2018, protection == "MPA") %>% 
  group_by(site) %>% 
  summarize(mean_size = mean(size_mm))

#nonMPA 2018 
lobster_mean_nonMPA_2018 <- lobster_protection %>% 
  filter(year == 2018, protection == "non-MPA") %>% 
  group_by(site) %>% 
  summarize(mean_size = mean(size_mm))

# MPA 2012
lobster_mean_MPA_2012 <- lobster_protection %>% 
  filter(year == 2012,
         protection == "MPA") %>% 
  group_by(site) %>% 
  summarize(mean_size = mean(size_mm))

#nonMPA 2012
lobster_mean_nonMPA_2012 <- lobster_protection %>% 
  filter(year == 2012,
         protection == "non-MPA") %>% 
  group_by(site) %>% 
  summarize(mean_size = mean(size_mm))

#### grouped by protection df's 
lobster_mean_12 <- lobster_protection %>%
  filter(year == 2012) %>% 
  group_by(protection) %>% 
  summarize(mean_size = mean(size_mm))

lobster_mean_18 <- lobster_protection %>%
  filter(year == 2018) %>% 
  group_by(protection) %>% 
  summarize(mean_size = mean(size_mm))


```

For 2012 observations, is there a significant difference in lobster size between MPA and non-MPA sites? 
```{r}
MPA_2012 <- lobster_protection %>% 
  filter(year == 2012,
         protection == "MPA")

nonMPA_2012 <- lobster_protection %>% 
  filter(year == 2012,
         protection == "non-MPA")

protection_2012 <- t.test(MPA_2012$size_mm, nonMPA_2012$size_mm)

```
Mean size (mm) for MPA sites (`r round(protection_2012$estimate[1],2)`) and non-MPA sites (`r round(protection_2012$estimate[2],2)`) in the year 2012 differ significantly (t(`r round(protection_2012$parameter,2)`) = (`r round(protection_2012$statistic,2)`, *p* = (`r round(protection_2012$p.value,2)`).

For 2018 observations, is there a significant difference in lobster size between MPA and non-MPA sites? 
```{r}

MPA_2018 <- lobster_protection %>% 
  filter(year == 2018, 
         protection == "MPA")

nonMPA_2018 <- lobster_protection %>% 
  filter(year == 2018, 
         protection == "non-MPA")

protection_2018 <- t.test(MPA_2018$size_mm, nonMPA_2018$size_mm)


```
Mean size (mm) for 2018 MPA sites (`r round(protection_2018$estimate[1],2)`) and non-MPA sites (`r round(protection_2018$estimate[2],2)`) in the year differ significantly (t(`r round(protection_2018$parameter,2)`) = (`r round(protection_2018$statistic,2)`, *p* = (`r round(protection_2018$p.value,2)`).

```{r}
#Representation of mean between MPA and non-MPA in 2018

status_all <- lobster_protection %>% 
  filter(year %in% c(2012:2018)) %>% 
  group_by(year, protection) %>% 
  mutate(size_avg = mean(size_mm))

ggplot(status_all, aes(size_mm, fill = protection))+
  geom_density(alpha=0.35) +
  scale_x_continuous(limits = c(30,200)) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9"))+
  geom_vline(aes(xintercept=size_avg,
                 color = protection),
             linetype = "dashed") +
  facet_wrap(~year) +
  labs(x = "Lobster Size (mm)",
       y = "Lobster Density",
       title = "Fig. 3: Change Lobster Size from 2012 to 2018",
       caption = "Figure 3 highlights the change in distribution among MPA and non-MPA sites between the years\n 2012 and 2018. The population means for all four distribution curves is indicated by a dashed line.") +
  theme_minimal()
```

For 2018 observations, is there a significant difference in lobster size between MPA and non-MPA sites? 
```{r}

MPA_2018 <- lobster_protection %>% 
  filter(year == 2018, 
         protection == "MPA")

nonMPA_2018 <- lobster_protection %>% 
  filter(year == 2018, 
         protection == "non-MPA")

protection_2018 <- t.test(MPA_2018$size_mm, nonMPA_2018$size_mm)


```
Mean size (mm) for 2018 MPA sites (`r round(protection_2018$estimate[1],2)`) and non-MPA sites (`r round(protection_2018$estimate[2],2)`) in the year differ significantly (t(`r round(protection_2018$parameter,2)`) = (`r round(protection_2018$statistic,2)`, *p* = (`r round(protection_2018$p.value,2)`).

For MPA sites only, is there a significant difference in lobsters observed in 2012 vs. 2018?
```{r}

MPA_all_2012 <- lobster_protection %>% 
  filter(year == 2012,
         protection == "MPA")

MPA_all_2018 <- lobster_protection %>% 
  filter(year == 2018,
         protection == "MPA")

protection_MPA_all <- t.test(MPA_all_2012$size_mm, MPA_all_2018$size_mm)

```
Mean size (mm) for 2012 MPA sites (`r round(protection_MPA_all$estimate[1],2)`) and 2018 MPA sites (`r round(protection_MPA_all$estimate[2],2)`) differ significantly (t(`r round(protection_MPA_all$parameter,2)`) = (`r round(protection_MPA_all$statistic,2)`, *p* = (`r round(protection_MPA_all$p.value,2)`).


For non-MPA sites only, is there a significant difference in lobsters observed in 2012 vs. 2018? 
```{r}

nonMPA_all_2012 <- lobster_protection %>% 
  filter(year == 2012,
         protection == "non-MPA")

nonMPA_all_2018 <- lobster_protection %>% 
  filter(year == 2018,
         protection == "non-MPA")

protection_nonMPA_all <- t.test(nonMPA_all_2012$size_mm, nonMPA_all_2018$size_mm)

```
Mean size (mm) for 2012 nonMPA sites (`r round(protection_nonMPA_all$estimate[1],2)`) and 2018 MPA sites (`r round(protection_nonMPA_all$estimate[2],2)`) do not differ significantly (t(`r round(protection_nonMPA_all$parameter,2)`) = (`r round(protection_nonMPA_all$statistic,2)`, *p* = (`r round(protection_nonMPA_all$p.value,2)`).



```{r}
# Creating a date frame with stats to make table
# include: avg, std dev & sample size for MPa and nonMPA 2012 and 2018 
mpa_stats <- lobster_protection %>%
  group_by(year, protection) %>% 
  summarize(
    mean_size = round(mean(size_mm, na.rm = TRUE), 2),
    standard_dev = round(sd(size_mm, na.rm = TRUE), 2),
    sample_size = n())

# Create a pretty table of the info!

mpa_stats_table <- mpa_stats %>% 
  kable(col.names = c("Year",
                      "Protection Status",
                      "Mean Lobster Size (mm)",
                      "Standard Deviation",
                      "Sample Size (n)")) %>% 
  kable_styling("striped") %>% 
  row_spec(c(1,3), color = "white", background = "#E69F00") %>%
  row_spec(c(2,4), color = "white", background = "#56B4E9")

mpa_stats_table

```

### Summary

A brief summary of the major findings (pick 3 - 4) from your mini-report. A bullet-pointed list is fine, but the findings should be well-written, responsible (don’t overstate your findings), and refer to outcomes (e.g. figures, tables) in the Results section

### References: 
https://www.wildlife.ca.gov/Conservation/Marine/MPAs



