---
title: "California Spiny Lobster Abundance"
author: "Zoe Duerksen-Salm and Audrey Nolan"
date: "11/18/2019"
output: html_document
---
Thoughts: 

- maybe title something like "Impact of Protection Status on California Spiny Lobster Abundance" ?
- move all of the figure caption/blurbs to before the graphs. Allison did this in her example report and i like the look of it. 
- How do we find a proper citation? I only found DOI. Do we need to include date data retrieved etc? **Under tables and graphs we could have a "source:" link that takes you to the original data source**
- for discussion of results from section C, do we want to remove the questions and just make a separate paragraph for each test, then talk about the main graph? Orrrrr talk about the main graph, and throw in the specific statitic values through the discussion. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

library(tidyverse)
library(tidyr)
library(janitor)
library(here)
library(plotly)
library(DT)
library(lubridate)
library(ggridges)
library(knitr)
library(kableExtra)

```

### Introduction 
A short (7 - 8 sentence) introduction that helps the reader understand the purpose/motivation of the research, and previews the content of the report.

This report explores California Spiny Lobster, *Panulirus interruptus*, populations from 2012-2018 at five sites off the coast of Santa Barbara, California. Spiny lobsters play a key role in the **[maintenance and health]** of giant kelp forests. Foraging on algae, snails, crabs, and sea urchins, spiny lobsters help **[manage kelp predator populations]**, thus sustaining a **[healthy, long-living population of kelp]**. The data - derived from **[]** - analyzes five study sites in the Santa Barbara area. Two of the sites, Naples and Isla Vista, became marine protected areas (MPA) in 2012. Under the Marine Life Protection Act, MPAs aim to preserve the diversity and abundance of marine life, the habitats they depend on, and the integrity of marine ecosystems(Cal Department of Fish and Wildlife). MPAs allow for.... 

### Data and Methods
A brief ‘Data and Methods’ section (5 - 6 sentences) summarizing the data and how it was analyzed. Briefly describe the variables being studied in this report. Include the types of statistical tests performed, significance level used, and software (with version) for analysis.

---

The data used for this report was obtained by reasearchers with Santa Barbara Coastal Long Term Ecological Research(SBC LTER) that began sampling in 2012 and **[remains]** ongoing. Lobster counts and lobster size (mm) are recorded by divers in four 300m^2 lobster transects. This data is collected annually in late summer at each of the five sites along the coast of the Santa Barbara Channel. Lobster size and abundance are two main varibles assessed within this report. 

Statistical tests were performed on the data including a Whelch Two Sample t-test with a significance level of p = 0.05. All statistical analyses and figures were prepared using R software version 3.6.1.

doi:10.6073/pasta/a593a675d644fdefb736750b291579a0

```{r}
#Read in lobster file + make it tidy

lobster_raw <- read_csv("lobster_abundance_sbc_lter.csv", na = "-99999")

lobster_tidy <- lobster_raw %>% 
  clean_names() %>%
  tidyr::uncount(count) %>% 
  select(year:size_mm)



```

### Results

Prepare finalized data visualizations and perform statistical analyses (as indicated) to do the following. You are expected to also add useful text that weaves the pieces together into a cohesive Results section (i.e., this should not just be a series of figures and without useful description & text flow between them). 

---

MPA sites established in 2012 - Isla Vista and Naples - are represented by orange coloration in all figures throughout this report. Non-MPA sites - Arroyo Quemado, Carpinteria, Mohawk - are indicated with blue coloration.

**[should this be in the methods section?]**

---

#### A: Changes in Lobster abundance (counts) by site.

The implementation of MPA status on certain study sites in 2012 produced an observable difference in both lobster size and abundance (Fig. 1). Over the six-year period post MPA **[implementation]**, the MPA sites - Naples and Isla Vista - **[exhibited]** an increase in the number of lobsters counted by researchers. Both sites had relatively low lobster numbers in 2012: Isla Vista was observed to have under 30, while less than 10 lobsters were found at Naples. By 2018, the Naples population had grown by almost 200 lobsters, and Isla Vista, with over 800 counted lobsters, contained highest abundance among the five sites. The non-MPA sites, Arroyo Quemada, Carpinteria, and Mohawk, show varying lobster abundance over the six-year period. Mohawk shows a decrease in abundance after the highest count at the site in 2014. This steady decline in abundance may be due to overharvesting during the peak abundance year and the lobster populations have **[struggled]** to recover.

```{r}
# creating dataframe labeling MPA vs nonMPA
lobster_abund <- lobster_tidy %>% 
  mutate(protection = ifelse(site %in% c("IVEE", "NAPL"), "MPA", "non-MPA"))

# Change site variable names from abreviations to full spelling
lobster_abund$site <- gsub("IVEE", "Isla Vista", lobster_abund$site)
lobster_abund$site <- gsub("NAPL", "Naples", lobster_abund$site)
lobster_abund$site <- gsub("AQUE", "Arroyo Quemado", lobster_abund$site)
lobster_abund$site <- gsub("CARP", "Carpinteria", lobster_abund$site)
lobster_abund$site <- gsub("MOHK", "Mohawk", lobster_abund$site)

```

```{r}
# Making site abundance graph

site_abundance_graph <- ggplot(lobster_abund, aes(x = year))+
  geom_histogram(binwidth = 1,
                 aes(fill = protection),
                 show.legend = FALSE) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9"))+
  facet_wrap(~site) +
  scale_x_continuous(limits = c(2011, 2019),
                     expand = c(0,0),
                     breaks = c(2012, 2014, 2016, 2018))+
  theme_light() +
  labs(x = "Year",
       y = "Lobster Count (n)",
       title = "Figure 1: Lobster Abundance Across Five Sites", 
       caption = "Figure 1 depicts the lobster abundance at each of the five sites. MPA sites are\n represented by orange coloration and non-MPA sites are shown by blue coloration.")

site_abundance_graph

```

---

#### B: Differences in lobster size distribution from 2012 to 2018

In 2012 when Naples and Isla Vista were designated as MPAs, the sites exhibited a relatively smaller distribution of lobster size. However, data from 2018 indicates that the populations at MPA sites shifted towards larger lobsters. Non-MPA sites show minimal change in the distrubution of lobster sizes over the six-year period. In fact, by 2018 the average lobster size at Naples and Isla Vista were 7.5% and 10.6% higher than in 2012 respectively. In comparison, mean lobster size at non-MPA only increased by 0.1-0.6% during the six-year-period. Figure 2 represents the shifts in the distribution of lobster size. The peaks indicate large quantities of small lobsters at MPA sites in 2012 as opposed to more sizable counts of larger lobsters at the same sites in 2018. In accordance with numerical results, the visual distributions of lobster size at non-MPA sites shows little variance between the two years. An increase in lobster size can be attributed to the **[designation]** of this area as an MPA, allowing for lobsters to mature without being harvested.

```{r}

lobster_12_18 <- lobster_tidy %>% 
  filter(year %in% c(2012, 2018))

lobster_all <- lobster_abund %>% 
  filter(year %in% c(2012, 2018))

lobster_size <- lobster_12_18 %>% 
  select(year, site, size_mm)

# Change site variable names from abreviations to full spelling
lobster_size$site <- gsub("IVEE", "Isla Vista", lobster_size$site)
lobster_size$site <- gsub("NAPL", "Naples", lobster_size$site)
lobster_size$site <- gsub("AQUE", "Arroyo Quemado", lobster_size$site)
lobster_size$site <- gsub("CARP", "Carpinteria", lobster_size$site)
lobster_size$site <- gsub("MOHK", "Mohawk", lobster_size$site)

```

```{r}
size_avg <- lobster_all %>% 
  group_by(site, year) %>% 
  summarise(median = median(size_mm),
            mean = mean(size_mm))

# Creating a graph to compare lobsters sizes in 2012 vs. 2018 

size_dist_by_site <- ggplot(lobster_all, aes(x = size_mm, y = site)) +
  geom_density_ridges(aes(fill = protection),
                      show.legend = FALSE)+
  scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
  facet_wrap(~year) +
  scale_x_continuous(limits = c(30, 135),
                     expand = c(0,0)) +
  theme_minimal()+
  labs(x = "Lobster Size (mm)",
       y = "Research Site",
       title = "Figure 2: Lobster Size Across Sites",
       caption = "Figure 2 illustrates the difference in lobsters size between 2012 and 2018 at all five sites.")

size_dist_by_site
```

---

#### C: Comparing mean lobster size at MPA vs. non-MPA sites in 2012 and 2018 

In text in your report, describe the outcome of the statistical hypothesis tests using in-line referencing (e.g. do not copy and paste values from the output). Remember: the p-value is not enough. You should also describe more meaningful metrics for comparison. Here are some things you might consider including in your description: the actual differences in mean size between groups, % changes, measure of uncertainty (e.g. confidence interval), etc.

```{r}
# For each year (2012 and 2018) consolidate the size observations into only two groups: MPA and non-MPA lobsters.

# Manipulate df to indicate MPA or non-MPA status. Filter df to include only 2012 + 2018 observations.

lobster_protection <- lobster_tidy %>% 
  mutate(protection = ifelse(site %in% c("IVEE", "NAPL"), "MPA", "non-MPA")) %>% 
  filter(year == c(2012, 2018)) %>% 
  select("year", "site", "size_mm", "protection")

#### wrangling to summarize mean of each 4 groups 

# MPA 2018
lobster_mean_MPA_2018 <- lobster_protection %>% 
  filter(year == 2018, protection == "MPA") %>% 
  group_by(site) %>% 
  summarize(mean_size = mean(size_mm))

#nonMPA 2018 
lobster_mean_nonMPA_2018 <- lobster_protection %>% 
  filter(year == 2018, protection == "non-MPA") %>% 
  group_by(site) %>% 
  summarize(mean_size = mean(size_mm))

# MPA 2012
lobster_mean_MPA_2012 <- lobster_protection %>% 
  filter(year == 2012,
         protection == "MPA") %>% 
  group_by(site) %>% 
  summarize(mean_size = mean(size_mm))

#nonMPA 2012
lobster_mean_nonMPA_2012 <- lobster_protection %>% 
  filter(year == 2012,
         protection == "non-MPA") %>% 
  group_by(site) %>% 
  summarize(mean_size = mean(size_mm))

#### grouped by protection df's 
lobster_mean_12 <- lobster_protection %>%
  filter(year == 2012) %>% 
  group_by(protection) %>% 
  summarize(mean_size = mean(size_mm))

lobster_mean_18 <- lobster_protection %>%
  filter(year == 2018) %>% 
  group_by(protection) %>% 
  summarize(mean_size = mean(size_mm))


```


```{r}
# Creating data frames to test for a significant difference in lobster size between MPA and non-MPA sites
MPA_2012 <- lobster_protection %>% 
  filter(year == 2012,
         protection == "MPA")

nonMPA_2012 <- lobster_protection %>% 
  filter(year == 2012,
         protection == "non-MPA")

protection_2012 <- t.test(MPA_2012$size_mm, nonMPA_2012$size_mm)

```
Mean size (mm) for MPA sites (`r round(protection_2012$estimate[1],2)`) and non-MPA sites (`r round(protection_2012$estimate[2],2)`) in the year 2012 differ significantly (t(`r round(protection_2012$parameter,2)`) = (`r round(protection_2012$statistic,2)`, *p* = (`r round(protection_2012$p.value,2)`).


```{r}
# Creating data frames to test for a significant difference in lobster size between MPA and non-MPA sites

MPA_2018 <- lobster_protection %>% 
  filter(year == 2018, 
         protection == "MPA")

nonMPA_2018 <- lobster_protection %>% 
  filter(year == 2018, 
         protection == "non-MPA")

protection_2018 <- t.test(MPA_2018$size_mm, nonMPA_2018$size_mm)


```
Mean size (mm) for 2018 MPA sites (`r round(protection_2018$estimate[1],2)`) and non-MPA sites (`r round(protection_2018$estimate[2],2)`) in the year differ significantly (t(`r round(protection_2018$parameter,2)`) = (`r round(protection_2018$statistic,2)`, *p* = (`r round(protection_2018$p.value,2)`).

```{r}
#Representation of mean between MPA and non-MPA in 2018

status_all <- lobster_protection %>% 
  filter(year %in% c(2012:2018)) %>% 
  group_by(year, protection) %>% 
  mutate(size_avg = mean(size_mm))

ggplot(status_all, aes(size_mm, fill = protection))+
  geom_density(alpha=0.35) +
  scale_x_continuous(limits = c(30,200)) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9"))+
  geom_vline(aes(xintercept=size_avg,
                 color = protection),
             linetype = "dashed") +
  facet_wrap(~year) +
  labs(x = "Lobster Size (mm)",
       y = "Lobster Density",
       title = "Figure 3: Change Lobster Size from 2012 to 2018",
       caption = "Figure 3 highlights the change in distribution among MPA and non-MPA sites between the years\n 2012 and 2018. The population means for all four distribution curves is indicated by a dashed line.") +
  theme_minimal()
```

For 2018 observations, is there a significant difference in lobster size between MPA and non-MPA sites? 
```{r}

MPA_2018 <- lobster_protection %>% 
  filter(year == 2018, 
         protection == "MPA")

nonMPA_2018 <- lobster_protection %>% 
  filter(year == 2018, 
         protection == "non-MPA")

protection_2018 <- t.test(MPA_2018$size_mm, nonMPA_2018$size_mm)


```
Mean size (mm) for 2018 MPA sites (`r round(protection_2018$estimate[1],2)`) and non-MPA sites (`r round(protection_2018$estimate[2],2)`) in the year differ significantly (t(`r round(protection_2018$parameter,2)`) = (`r round(protection_2018$statistic,2)`, *p* = (`r round(protection_2018$p.value,2)`).

For MPA sites only, is there a significant difference in lobsters observed in 2012 vs. 2018?
```{r}

MPA_all_2012 <- lobster_protection %>% 
  filter(year == 2012,
         protection == "MPA")

MPA_all_2018 <- lobster_protection %>% 
  filter(year == 2018,
         protection == "MPA")

protection_MPA_all <- t.test(MPA_all_2012$size_mm, MPA_all_2018$size_mm)

```
Mean size (mm) for 2012 MPA sites (`r round(protection_MPA_all$estimate[1],2)`) and 2018 MPA sites (`r round(protection_MPA_all$estimate[2],2)`) differ significantly (t(`r round(protection_MPA_all$parameter,2)`) = (`r round(protection_MPA_all$statistic,2)`, *p* = (`r round(protection_MPA_all$p.value,2)`).


For non-MPA sites only, is there a significant difference in lobsters observed in 2012 vs. 2018? 
```{r}

nonMPA_all_2012 <- lobster_protection %>% 
  filter(year == 2012,
         protection == "non-MPA")

nonMPA_all_2018 <- lobster_protection %>% 
  filter(year == 2018,
         protection == "non-MPA")

protection_nonMPA_all <- t.test(nonMPA_all_2012$size_mm, nonMPA_all_2018$size_mm)

```
Mean size (mm) for 2012 nonMPA sites (`r round(protection_nonMPA_all$estimate[1],2)`) and 2018 MPA sites (`r round(protection_nonMPA_all$estimate[2],2)`) do not differ significantly (t(`r round(protection_nonMPA_all$parameter,2)`) = (`r round(protection_nonMPA_all$statistic,2)`, *p* = (`r round(protection_nonMPA_all$p.value,2)`).



```{r}
# Creating a date frame with stats to make table
# include: avg, std dev & sample size for MPa and nonMPA 2012 and 2018 
mpa_stats <- lobster_protection %>%
  group_by(year, protection) %>% 
  summarize(
    mean_size = round(mean(size_mm, na.rm = TRUE), 2),
    standard_dev = round(sd(size_mm, na.rm = TRUE), 2),
    sample_size = n())

# Create a pretty table of the info!

mpa_stats_table <- mpa_stats %>% 
  kable(col.names = c("Year",
                      "Protection Status",
                      "Mean Lobster Size (mm)",
                      "Standard Deviation",
                      "Sample Size (n)")) %>% 
  kable_styling("striped") %>% 
  row_spec(c(1,3), color = "white", background = "#E69F00") %>%
  row_spec(c(2,4), color = "white", background = "#56B4E9")

mpa_stats_table

```

### Summary

A brief summary of the major findings (pick 3 - 4) from your mini-report. A bullet-pointed list is fine, but the findings should be well-written, responsible (don’t overstate your findings), and refer to outcomes (e.g. figures, tables) in the Results section

### References: 
https://www.wildlife.ca.gov/Conservation/Marine/MPAs



