---
title: "ESM 206: Assignment 4"
author: "Zoe Duerksen-Salm & Audrey Nolan"
date: "11/6/2019"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
```

```{r}
# Attach libraries

library(tidyverse)
library(tidyr)
library(janitor)
library(here)
library(plotly)
library(DT)
library(lubridate)
library(ggridges)
library(knitr)
library(kableExtra)

```
### Introduction 
A short (7 - 8 sentence) introduction that helps the reader understand the purpose/motivation of the research, and previews the content of the report.

### Data and Methods
A brief ‘Data and Methods’ section (5 - 6 sentences) summarizing the data and how it was analyzed. Briefly describe the variables being studied in this report. Include the types of statistical tests performed, significance level used, and software (with version) for analysis.
```{r}
#Read in lobster file + make it tidy

lobster_raw <- read_csv("lobster_abundance_sbc_lter.csv", na = "-99999")

lobster_tidy <- lobster_raw %>% 
  clean_names() %>%
  tidyr::uncount(count) %>% 
  select(year:size_mm)
```

### Results
#### Results A: Changes in Lobster abundance (counts) by site.
Add a figure caption below the graph.

In text (above or below the graph for Results A), describe general / interesting trends that you observe for lobster abundance at the 5 sites.

```{r}
lobster_abund <- lobster_tidy %>% 
  mutate(protection = ifelse(site %in% c("IVEE", "NAPL"), "MPA", "non-MPA"))

# Change site variable names from abreviations to full spelling
lobster_abund$site <- gsub("IVEE", "Isla Vista", lobster_abund$site)
lobster_abund$site <- gsub("NAPL", "Naples", lobster_abund$site)
lobster_abund$site <- gsub("AQUE", "Arroyo Quemado", lobster_abund$site)
lobster_abund$site <- gsub("CARP", "Carpinteria", lobster_abund$site)
lobster_abund$site <- gsub("MOHK", "Mohawk", lobster_abund$site)

# Make a graph

site_abundance_graph <- ggplot(lobster_abund, aes(x = year))+
  geom_histogram(binwidth = 1,
                 aes(fill = protection),
                 show.legend = FALSE) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9"))+
  facet_wrap(~site) +
  theme_light() +
  labs(x = "Year",
       y = "Lobster Count",
       title = "Lobster Abundance Across Five Sites(Duerksen-Salm & Nolan)", 
       caption = "YAY CAPTION HERE")

site_abundance_graph

## I dont like this caption but we need to talk about what the shaded area means / if the shaded area is necessary? 
(site_abundance_graph)

```

#### Results B: 

Visually explore lobster size distribution shifts
Compare lobster sizes in 2012 and 2018

```{r}
lobster_12_18 <- lobster_tidy %>% 
  filter(year %in% c(2012, 2018))

lobster_all <- lobster_abund %>% 
  filter(year %in% c(2012, 2018))
```

Visually explore the size distributions for lobsters observed at the 5 different sites in 2012, and in 2018
(i.e. you will have a total of 10 size distributions shown: for each of the 5 sites, one for 2012 and one for 2018).
```{r}
lobster_size <- lobster_12_18 %>% 
  select(year, site, size_mm)

# Change site variable names from abreviations to full spelling
lobster_size$site <- gsub("IVEE", "Isla Vista", lobster_size$site)
lobster_size$site <- gsub("NAPL", "Naples", lobster_size$site)
lobster_size$site <- gsub("AQUE", "Arroyo Quemado", lobster_size$site)
lobster_size$site <- gsub("CARP", "Carpinteria", lobster_size$site)
lobster_size$site <- gsub("MOHK", "Mohawk", lobster_size$site)
```


```{r}

# Change in distribution of lobster size
# Change in lobster size by site

ggplot(lobster_all, aes(x = size_mm, y = site)) +
  geom_density_ridges(aes(fill = protection),
                      show.legend = FALSE)+
  scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
  facet_wrap(~year) +
  scale_x_continuous(limits = c(30, 135),
                     expand = c(0,0)) +
  theme_light()+
  labs(x = "Lobster Size (mm)",
       y = "Site",
       title = "Change in lobster size by site", caption = "NEED CAPTION BELOW GRAPH")


```

In text, describe general / interesting trends that you see in community size structure at different sites (again, considering MPA vs. non-MPA status) for lobsters between 2012 and 2018.

---

Things that really didn't work but might be useful later...

```{r, eval = FALSE}
ggplot(lobster_size, aes(x = size_mm)) +
  geom_histogram(aes(fill = year,
                 show.legend = FALSE)) +
  facet_wrap(~site)+
  facet_wrap(~year)+
  labs(title = "Temperature Histogram", 
       y = "Lobster Count", 
       x = "Lobster Size (mm)") +
  theme_light()
```

```{r, eval = FALSE}
lobster_count <- lobster_raw %>% 
  clean_names() %>%
  group_by(year)

lobster_3a <- lobster_tidy %>% 
  select(date, site) %>% 
  group_by(site) %>% 
  count(date)

ggplot(lobster_3a, aes(x = date, y = n))+
  geom_line()+
  facet_wrap(~site)
```

#### Results C: Compare mean lobster sizes at MPA vs. non-MPA sites in 2012 and 2018 
- MPAs: IV and Naples (establish 2012)

For each year (2012 and 2018) consolidate the size observations into only two groups: MPA and non-MPA lobsters.

```{r}
# Manipulate df to indicate MPA or non-MPA status. Filter df to include only 2012 + 2018 observations.

lobster_protection <- lobster_tidy %>% 
  mutate(protection = ifelse(site %in% c("IVEE", "NAPL"), "MPA", "non-MPA")) %>% 
  filter(year == c(2012, 2018)) %>% 
  select("year", "site", "size_mm", "protection")



#### wrangling to summarize mean of each 4 groups 
## ***some of these df's have no results but im too braindead to explore why, will assess tmrw

# MPA 2018
lobster_mean_MPA_2018 <- lobster_protection %>% 
  filter(year == 2018, protection == "MPA") %>% 
  group_by(site) %>% 
  summarize(mean_size = mean(size_mm))

#nonMPA 2018 
lobster_mean_nonMPA_2018 <- lobster_protection %>% 
  filter(year == 2018, protection == "non-MPA") %>% 
  group_by(site) %>% 
  summarize(mean_size = mean(size_mm))

# MPA 2012
lobster_mean_MPA_2012 <- lobster_protection %>% 
  filter(year == 2012,
         protection == "MPA") %>% 
  group_by(site) %>% 
  summarize(mean_size = mean(size_mm))

#nonMPA 2012
lobster_mean_nonMPA_2012 <- lobster_protection %>% 
  filter(year == 2012,
         protection == "non-MPA") %>% 
  group_by(site) %>% 
  summarize(mean_size = mean(size_mm))

lobster_mean_12 <- lobster_protection %>%
  filter(year == 2012) %>% 
  group_by(protection) %>% 
  summarize(mean_size = mean(size_mm),
            )

lobster_mean_18 <- lobster_protection %>%
  filter(year == 2018) %>% 
  group_by(protection) %>% 
  summarize(mean_size = mean(size_mm))


```

```{r, eval = FALSE}
# Creating graph to visualize MPA vs nonMPa of 2012 to 2018 
## Trying to create a visualization of the increase mean lobster size in MPA zones from 2012--2018

ggplot() +
  geom_col(data = lobster_mean_12, aes(x = protection, y = mean_size), fill = "blue") +
  geom_col(data = lobster_mean_18, aes(x = protection, y = mean_size), fill = "green")

```

For 2012 observations, is there a significant difference in lobster size between MPA and non-MPA sites? 
```{r}
MPA_2012 <- lobster_protection %>% 
  filter(year == 2012,
         protection == "MPA")

nonMPA_2012 <- lobster_protection %>% 
  filter(year == 2012,
         protection == "non-MPA")

protection_2012 <- t.test(MPA_2012$size_mm, nonMPA_2012$size_mm)

```

Mean size (mm) for MPA sites (`r round(protection_2012$estimate[1],2)`) and non-MPA sites (`r round(protection_2012$estimate[2],2)`) in the year 2012 differ significantly (t(`r round(protection_2012$parameter,2)`) = (`r round(protection_2012$statistic,2)`, *p* = (`r round(protection_2012$p.value,2)`).

```{r}
#Representation of mean between MPA and non-MPA in 2012
status_12 <- lobster_protection %>%
  filter(year == 2012) %>% 
  group_by(protection) %>% 
  mutate(size_avg = mean(size_mm))

ggplot(status_12, aes(size_mm, fill = protection))+
  geom_density(alpha=0.3)+
  geom_vline(aes(xintercept=size_avg,
                 color = protection),
             linetype = "dashed")+
  scale_fill_manual(values = c("#E69F00", "#56B4E9"))+
  theme_minimal()
```


For 2018 observations, is there a significant difference in lobster size between MPA and non-MPA sites? 
```{r}
MPA_2018 <- lobster_protection %>% 
  filter(year == 2018, 
         protection == "MPA")

nonMPA_2018 <- lobster_protection %>% 
  filter(year == 2018, 
         protection == "non-MPA")

protection_2018 <- t.test(MPA_2018$size_mm, nonMPA_2018$size_mm)

```

Mean size (mm) for 2018 MPA sites (`r round(protection_2018$estimate[1],2)`) and non-MPA sites (`r round(protection_2018$estimate[2],2)`) in the year differ significantly (t(`r round(protection_2018$parameter,2)`) = (`r round(protection_2018$statistic,2)`, *p* = (`r round(protection_2018$p.value,2)`).

```{r}
#Representation of mean between MPA and non-MPA in 2018
status_18 <- lobster_protection %>%
  filter(year == 2018) %>%
  group_by(protection) %>% 
  mutate(size_avg = mean(size_mm))

ggplot(status_18, aes(size_mm, fill = protection))+
  geom_density(alpha=0.2)+
  geom_vline(aes(xintercept=size_avg,
                 color = protection),
             linetype = "dashed")

status_all <- lobster_protection %>% 
  filter(year %in% c(2012:2018)) %>% 
  group_by(year) %>% 
  mutate(size_avg = mean(size_mm))

ggplot(status_all, aes(size_mm, fill = protection))+
  geom_density(alpha=0.3) +
  scale_x_continuous(limits = c(30,200)) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9"))+
  geom_vline(aes(xintercept=size_avg,
                 color = protection),
             linetype = "dashed") +
  facet_wrap(~year) +
  theme_minimal()

```


For MPA sites only, is there a significant difference in lobsters observed in 2012 vs. 2018?
```{r}
MPA_all_2012 <- lobster_protection %>% 
  filter(year == 2012,
         protection == "MPA")

MPA_all_2018 <- lobster_protection %>% 
  filter(year == 2018,
         protection == "MPA")

protection_MPA_all <- t.test(MPA_all_2012$size_mm, MPA_all_2018$size_mm)

```
Mean size (mm) for 2012 MPA sites (`r round(protection_MPA_all$estimate[1],2)`) and 2018 MPA sites (`r round(protection_MPA_all$estimate[2],2)`) differ significantly (t(`r round(protection_MPA_all$parameter,2)`) = (`r round(protection_MPA_all$statistic,2)`, *p* = (`r round(protection_MPA_all$p.value,2)`).



For non-MPA sites only, is there a significant difference in lobsters observed in 2012 vs. 2018? 
```{r}

nonMPA_all_2012 <- lobster_protection %>% 
  filter(year == 2012,
         protection == "non-MPA")

nonMPA_all_2018 <- lobster_protection %>% 
  filter(year == 2018,
         protection == "non-MPA")

protection_nonMPA_all <- t.test(nonMPA_all_2012$size_mm, nonMPA_all_2018$size_mm)

```
Mean size (mm) for 2012 nonMPA sites (`r round(protection_nonMPA_all$estimate[1],2)`) and 2018 MPA sites (`r round(protection_nonMPA_all$estimate[2],2)`) do not differ significantly (t(`r round(protection_nonMPA_all$parameter,2)`) = (`r round(protection_nonMPA_all$statistic,2)`, *p* = (`r round(protection_nonMPA_all$p.value,2)`).
 

```{r}
# Creating a date frame with stats to make table
# include: avg, std dev & sample size for MPa and nonMPA 2012 and 2018 
mpa_stats <- lobster_protection %>%
  group_by(year, protection) %>% 
  summarize(
    mean_size = round(mean(size_mm, na.rm = TRUE), 2),
    standard_dev = round(sd(size_mm, na.rm = TRUE), 2),
    sample_size = n())

# Create a pretty table of the info!

mpa_stats_table <- mpa_stats %>% 
  kable(col.names = c("Year",
                      "Protection Status",
                      "Mean Lobster Size (mm)",
                      "Standard Deviation",
                      "Sample Size (n)")) %>% 
  kable_styling("striped") %>% 
  row_spec(c(1,3), color = "white", background = "#E69F00") %>%
  row_spec(c(2,4), color = "white", background = "#56B4E9")

mpa_stats_table

```





## To do: 
- get dylan's code to combine stats and make a table 
- code for ^^ is in lab 5 ?? 
```{r, eval=FALSE}

datatable(protection_2012, protection_2018, protection_MPA_all, protection_nonMPA_all)

```

